Fast and adaptive cointegration based model for forecasting financial time series

### 2017-12-08 slides/defense-slides.tex
New theme and heteroscedasticity def

\documentclass{beamer}
\usetheme{CambridgeUS}
%\usecolortheme{lily}
%\usetheme{metropolis}
%\usepackage[spanish,activeacute]{babel}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{color}
\usepackage{tikz}
\usepackage{array}
%\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{booktabs}
\usetikzlibrary{shapes,arrows}
%\usepackage{multicol}
\usepackage{hhline}
\usepackage{pgf}
\usepackage{algorithm} 
\usepackage{algorithmic}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\def\R{{\mathbb R }}
\def\N{{\mathbb N }}

%   \definecolor{myblue1}{RGB}{35,119,189}
%    \definecolor{myblue2}{RGB}{95,179,238}
%    \definecolor{myblue3}{RGB}{129,168,207}
%    \definecolor{myblue4}{RGB}{26,89,142}
%
%    \setbeamercolor*{structure}{fg=myblue1,bg=blue}
%    \setbeamercolor*{palette primary}{use=structure,fg=white,bg=structure.fg}
%    \setbeamercolor*{palette secondary}{use=structure,fg=white,bg=structure.fg!75!black}
%    \setbeamercolor*{palette tertiary}{use=structure,fg=white,bg=structure.fg!50!black}
%    \setbeamercolor*{palette quaternary}{fg=black,bg=white}
%
%    \setbeamercolor*{item projected}{fg=red,bg=myblue3!80}
%    \setbeamercolor*{block title example}{fg=white,bg=myblue4}
%    \setbeamercolor*{frametitle}{fg=black}
%
%    \setbeamertemplate{blocks}[rounded][shadow=true]
%
%    \makeatletter
%    \pgfdeclarehorizontalshading[frametitle.bg,frametitle right.bg]{beamer@frametitleshade}{\paperheight}{%
%      color(0pt)=(myblue2);
%      color(\paperwidth)=(white)}
%
%    \defbeamertemplate*{footline}{mysplit theme}
%    {%
%      \leavevmode%
%      \hbox{\begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm plus1fill,rightskip=.3cm]{author in head/foot}%
%        \usebeamerfont{author in head/foot}\insertshortauthor
%      \end{beamercolorbox}%
%      \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
%        \usebeamerfont{title in head/foot}\insertshorttitle\hfill
%        \insertframenumber/\inserttotalframenumber\hspace*{0.5em}
%      \end{beamercolorbox}}%
%      \vskip0pt%
%    }
%    \makeatother

%\usebackgroundtemplate{%             declare it
%\tikz[overlay,remember picture] \node[opacity=0.2, at=(current page.center)] {
%   \includegraphics[height=\paperheight,width=\paperwidth]{img/background}};
%}

\author[Paola Arce]{Paola Arce \\ Departamento de Inform\'atica, UTFSM \\ Thesis advisor: Luis Salinas}
\date{December 15, 2017}
\title[Thesis defense]{Online learning methods for financial time series forecasting}
\subtitle{PhD Thesis Defense}

@@

\begin{frame}
\frametitle{HFT Revenue}
However, the revenues have fallen dramatically mainly because of more regulation, increasing cost of HFT infrastructure and competition.
\begin{figure}
\includegraphics[width=0.7\paperwidth]{img/HFTmarket2}
\caption{Revenue in the US. Source: TABB group}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{High Frequency data}
High frequency time series are commonly:
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{itemize}
\item Non-stationary
\item Irregularly spaced over time.
%\item Heterogeneity
\item Heteroscedastic
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{figure}
\includegraphics[width=0.35\paperwidth]{img/forecast}
\end{figure}
\end{column}
\end{columns}
\vspace{5mm}
This imply that classical statistical models can't be used. Therefore, financial time series forecasting is still a challenge.
\end{frame}

\begin{frame}
\frametitle{Stationarity}
\begin{block}{Strong definition}
A strictly stationary time series $y_t \in \mathbb{R}$, $\forall t \in \mathbb{Z}$ is one for which the probabilistic behaviour of every collection of values $\{y_{t_1},y_{t_2},\dots,y_{t_L}\}$ is identical
to that of the time shifted set, more precisely:
\small

\begin{equation*}
{\color{blue}
P\{y_{t_1} \leq
c_1,..,y_{t_L} \leq c_L\} = P\{y_{t_1+h} \leq c_1,..,y_{t_L+h} \leq c_L\}\enspace }\forall L \in \mathbb{N}, \forall h \in \mathbb{Z}
\end{equation*}
\noindent where $c_1,\dots,c_L$ are constants.
\end{block}
\end{frame}

@@

\begin{frame}
\frametitle{Heteroscedastic}
%HASTA AQUI
\end{frame}

>>>>

\documentclass{beamer}
%\usepackage[spanish,activeacute]{babel}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{color}
\usepackage{tikz}
\usepackage{array}
%\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{booktabs}
\usetikzlibrary{shapes,arrows}
%\usepackage{multicol}
\usepackage{hhline}
\usepackage{pgf}
\usepackage{algorithm} 
\usepackage{algorithmic}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\def\R{{\mathbb R }}
\def\N{{\mathbb N }}

\usetheme{metropolis}

\definecolor{BlueTOL}{HTML}{222255}
\definecolor{BrownTOL}{HTML}{666633}
\definecolor{GreenTOL}{HTML}{225522}
\setbeamercolor{normal text}{fg=BlueTOL,bg=white}
\setbeamercolor{alerted text}{fg=BrownTOL}
\setbeamercolor{example text}{fg=GreenTOL}

\metroset{block=fill}

\setbeamercolor{block title alerted}{use=alerted text,
    fg=alerted text.fg,
    bg=alerted text.bg!80!alerted text.fg}
\setbeamercolor{block body alerted}{use={block title alerted, alerted text},
    fg=alerted text.fg,
    bg=block title alerted.bg!50!alerted text.bg}
\setbeamercolor{block title example}{use=example text,
    fg=example text.fg,
    bg=example text.bg!80!example text.fg}
\setbeamercolor{block body example}{use={block title example, example text},
    fg=example text.fg,
    bg=block title example.bg!50!example text.bg}

%\author[Paola Arce]{Paola Arce \\ Departamento de Inform\'atica, UTFSM \\ Thesis advisor: Luis Salinas}
%\date{December 15, 2017}
%\title[Thesis defense]{Online learning methods for financial time series forecasting}
%\subtitle{PhD Thesis Defense}
%
%\titlegraphic{
%    \tikz[overlay,remember picture]
%        \node[at=(current page.south east), anchor=south east] {
%            \includegraphics[height=1.5cm]{img/DIT}
%        };
%}

\author[Paola Arce]{Paola Arce \\ Departamento de Inform\'atica, UTFSM \\ Thesis advisor: Luis Salinas}
\date{December 11, 2017}
\title[Thesis defense]{Fast and adaptive cointegration based model for forecasting financial time series}
\subtitle{Seminario Inform\'atica CCTVal}

@@

\begin{frame}
\frametitle{HFT Revenue}
However, the revenues have fallen dramatically mainly because of more regulation, increasing cost of HFT infrastructure and more competitive algorithms.
\begin{figure}
\includegraphics[width=0.7\paperwidth]{img/HFTmarket2}
\caption{Revenue in the US. Source: TABB group}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{High Frequency data}
High frequency time series are commonly:
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{itemize}
\item Non-stationary
\item Irregularly spaced over time.
%\item Heterogeneity
\item Heteroscedastic
%\item Volatility
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{figure}
\includegraphics[width=0.35\paperwidth]{img/forecast}
\end{figure}
\end{column}
\end{columns}
\vspace{5mm}
This imply that classical statistical models can't be used. Therefore, financial time series forecasting is still a challenge.
\end{frame}

\begin{frame}
\frametitle{Stationarity}
\begin{block}{Strong definition}
A {\color{red}strictly stationary time series} $y_t \in \mathbb{R}$, $\forall t \in \mathbb{Z}$ is one for which the probabilistic behaviour of every collection of values $\{y_{t_1},y_{t_2},\dots,y_{t_L}\}$ is identical
to that of the time shifted set, more precisely:
\begin{equation*}
{\color{blue}
P\{y_{t_1} \leq
c_1,..,y_{t_L} \leq c_L\} = P\{y_{t_1+h} \leq c_1,..,y_{t_L+h} \leq c_L\}\enspace }
\end{equation*}
\noindent $\forall L \in \mathbb{N}, \forall h \in \mathbb{Z}$ where $c_1,\dots,c_L$ are constants.
\end{block}
\end{frame}

@@

\begin{frame}
\frametitle{Heteroscedastic}
Conditional heteroskedasticity identifies non-constant volatility.
Volatility gives a measure of uncertainty about future returns. Many factors affect volatility such as: historical prices, trends, monetary policies and news.
\begin{figure}
\includegraphics[width=0.8\paperwidth]{img/vix}
\end{figure}
\end{frame}
